Correlation tests

### 2025-04-07 analysis/manuscript.tex
First complete draft, to be shared

Transformative agreements have emerged as a much-discussed licensing
model for the transition from subscription-based publishing to full open
access in the last ten years. While these agreements can vary
substantially, increasing coordination exists in standardising the
negotiation, administration, and evaluation of these agreements.
Simultaneously, efforts to improve transparency through open data about
contracts and payments are growing. However, this information remains
only partially available through bibliometric databases, limiting
large-scale studies about the impact of transformative agreements on the
transition of hybrid journals to full open access through increased open
access uptake. This study compares findings from hoaddata, an openly
available, continuously updated dataset on hybrid open access, built on
open data from Crossref, OpenAlex and the cOAlition S Journal Checler
Tool, with the established bibliometric databases Web of Science and
Scopus. The purpose of this comparison is not only to evaluate the
suitability of open metadata for measuring hybrid open access but also
to inform bibliometric research and practice by highlighting the
strengths and weaknesses of each data source.

Since their introduction, transformative agreements have sparked
considerable academic debate. Whilst standardisation activities,
particularly through the OA2020 and ESAC (Efficiency and Standards for
Article Charges) initiatives, have led to improved workflows between
library consortia and publishers, as well as increased open access
publishing, the initial focus on a few large publishers by national
consortia from high-income countries has drawn criticism. Critics have
raised concerns about market concentration, the lack of journal flipping
to full open access models, reduced competition in the publishing
market, and wider gaps between well-resourced and under-resourced
institutions and researchers.

For instance, studying over 700 agreements, Jahn (2025) observed a
strong growth in open access publishing to 15\% in 2022 due to the
varying implementation of transformative agreements across countries,
predominantly driven by three commercial publishers: Elsevier, Springer
Nature, and Wiley. This aligns with the rapid proliferation of
transformative agreements in recent years. Previously, only a minimal
proportion of articles were made open access in hybrid journals
(Piwowar, 2018; Laakso, 2016).

Evaluations conducted by library consortia confirm these trends but draw
different conclusions. For instance, the British Joint Information
Systems Committee's evaluation (Brayman, Devenney et al., 2024) noted
transformative agreements' significant impact on national open access
growth but limited effect on global open access transition, suggesting a
need to reevaluate their strategy. Similar policy reconsiderations
emerged in Norway and Sweden (Holden, Skoie et al., 2023; Widding,
2024). Meanwhile, cOAlition S recommended ending financial support for
these agreements in 2024, but still considers hybrid open access
compliant with funder open access policies. Conversely, Germany's DEAL
consortium extended agreements with major publishers through 2028, while
Colombia established Latin America's first transformative agreements
(Muñoz-Vélez, Pallares et al., 2024). A recent Max Planck Digital
Library study demonstrates how cost savings can be achieved through this
model and how it can be further developed to achieve full open access
while controlling expenditure.

Open databases like OpenAlex, which build upon Crossref, particularly
benefit from these developments. These resources have gained
considerable attention in research evaluation due to their inclusivity,
openness, and comprehensive coverage of publication metadata. Recent
studies {[}REFERENCE NEEDED{]} have demonstrated that open bibliometric
databases increasingly match or exceed the coverage of traditional
commercial databases in many disciplines, whilst providing improved
access to standardised and interlinked open research information.

Against this background, this study evaluates the suitability of open
metadata to measure the impact of transformative agreements on hybrid
open access. Only a few studies have attempted this analysis, primarily
due to data limitations and the need to combine multiple sources.
Transformative agreements present unique measurement challenges because
of their complex workflow requirements, which are not fully reflected in
bibliometric databases. These challenges include varying journal
portfolios across agreements, invoicing procedures specific to
submitting authors, and article caps that limit open access publishing
within agreements.

The present study addresses these gaps by replicating and applying
consistent methodologies across Web of Science, Scopus, and open data
sources. This comparative approach enables the research to overcome the
limitations of any single source whilst systematically assessing the
strengths and weaknesses of different bibliometric data sources for
monitoring transformative agreement outcomes. By triangulating findings
across these platforms, the study aims to provide a more comprehensive
understanding of how transformative agreements influence hybrid open
access publishing patterns within the ongoing scholarly publishing
transition.

@@

@@

This study of over 13,000 hybrid journals shows a rise in open access
due to transformative agreements between 2019 and 2023, although most
articles remained paywalled. While transformative agreements accounted
for the majority of open access, many articles continue to become open
through the payment of individual publication fees. Hybrid open access
and transformative agreements remain concentrated among a small group of
large commercial publishers, with high-income European
countries---alongside South Africa---showing high adoption rates. In
contrast, the three most productive countries, China, the United States,
and India, were substantially ranked lower in transformative agreement
adoption. Open questions remain whether this uneven distribution
reflects temporary implementation gaps, inherent inequities in the
transformative agreement model, or countries deliberately refraining
from such agreements.

The findings were consistent across the investigated open data source
hoaddata, derived from Crossref and OpenAlex, and the established
proprietary bibliometric databases Scopus and Web of Science. Notably,
the results show strong correlations by country affiliation, aligning
with previous studies (Akbaritabar et al., 2024; Alperin et al., 2024)
and rankings (Eck et al., 2024), which supports using open metadata for
large-scale analyses of hybrid open access. However, the observed
differences in journal coverage and metadata availability warrant
discussion, affecting not only open data sources but also proprietary
databases when used in isolation.

The coverage analysis reveals that hybrid journals are well indexed in
all three data sources, particularly in terms of article coverage.
Differences emerge for journals targeting practitioners or local
non-English language communities, with many such titles indexed
exclusively in Crossref and Scopus. Using the open index Crossref as a
bibliometric database in hoaddata demonstrated particular strength in
identifying newly established hybrid journals, a notable finding given
that transformative agreements primarily target existing
subscription-based journals.

The landscape of hybrid journal publishing differs markedly from that of
fully open access journals. Comparing the coverage of OpenAlex, Scopus
and Web of Science, Simard et al. (2024) indicate that only half of the
fully open access journals listed in the DOAJ are also indexed in Scopus
and Web of Science. Notably, journals that charge no publication fees
(``diamond journals'') are absent from the selective Web of Science,
which reinforces existing disparities in the indexing of
underrepresented research communities and regions in selective
bibliometric databases (ref).

A frequently reported limitation of studying open access with less
selective databases is the lack of corresponding authorship information
(Fraser et al., 2023; Haucap et al., 2021; Shu \& Larivière, 2023).
However, this analysis demonstrates that indicators based on first
authors, which have often been used as a proxy for determining open
access funding, and corresponding authors show a high level of
correlation, reflecting disciplinary norms in scholarly publishing with
regard to contributorship and author roles and positions (Larivière et
al., 2016). First authors typically conduct the main research underlying
a paper, while the corresponding author often supervises the research
(Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly, measures
based on first or corresponding authorships are strongly correlated,
suggesting that these authors share the same country affiliation.
Moreover, in most cases, the first author is identical to the
corresponding author (Chinchilla-Rodríguez et al., 2024). Thus, first
authorship can reliably be used in large-scale studies to assess the
impact of transformative agreements.

Despite this correlation, the study observed a recent slight increase in
open access articles by corresponding authors over first authors across
Springer Nature hybrid journals prompts questions about how
institutional open access sponsorship practices influence author roles
and assignments within co-author teams, especially as funding
opportunities vary (Gumpenberger et al., 2018). Previous research has
highlighted how institutionalized bibliometric practices can affect the
valuation of authorship positions (Helgesson, 2020), suggesting the need
to monitor potential influences on authorships (Maddi \& Silva, 2024).

@@

The database comparison revealed important discrepancies in open access
evidence. Crucially, not all publishers deposit CC licences with
Crossref, a limitation that becomes apparent when contrasting Crossref
with Unpaywall's data in Scopus and Web of Science. While Unpaywall can
detect such gaps by parsing publishers' websites for open access
licences, it nevertheless missed a substantial number of CC-licensed
open access articles from Wiley journals indexed in Crossref, likely due
to parsing errors on the publisher's website. This resulted in fewer
open access articles being recorded in Scopus and Web of Science.
Further inconsistencies emerged between Scopus and Web of Science,
despite both relying on Unpaywall: Web of Science erroneously labels
Elsevier's delayed open access as hybrid, whereas Scopus correctly
categorises it. Notably, Scopus and Crossref showed greater alignment
than observed in a related comparison between Scopus and OpenAlex
(Alperin et al., 2024). This likely reflects the use of a curated list
to identify hybrid open access journals, rather than relying solely on
article-level tags. Given these variations, which may compromise
comparability, research and monitoring exercises should avoid reliance
on a single source. Instead, the selection of open access evidence
should be cross-verified using multiple sources (Huang et al., 2020).
Incorporating expert-curated journal lists can also help prevent
misclassifications based on business models (Visser et al., 2021).

It is important to note that the study's estimates of articles from
institutions involved in transformative agreements are approximations
due to a lack of access to invoice data, which is not usually shared by
library consortia and publishers. Furthermore, the data sources about
transformative agreements including the cOAlition S Journal Checker Tool
and the ESAC registry are a voluntary effort, crowd-sourced from various
consortia. However, recent validation with Dutch Research Information
demonstrates the reliability of such an open approach for the assessment
of articles under transformative agreements (Jonge et al., 2025). With
enhanced and streamlined metadata compliance, attributable to evolving
standards and initiatives to support negotiations with publishers,
particularly through the communities of the ESAC initiative and
Barcelona Declaration supporters, the situation is likely to improve. In
conclusion, this study has shown that open scholarly data are suitable
for analysing the transition from subscription-based publishing to full
open access.

@@

@@

\bibitem[\citeproctext]{ref-Birkle_2020}
Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). Web of
science as a data source for research on scientific and scholarly
activity. \emph{Quantitative Science Studies}, \emph{1}(1), 363--376.
\url{https://doi.org/10.1162/qss_a_00018}

@@

@@

@@

@@

@@

@@

\bibitem[\citeproctext]{ref-Huang_2020}
Huang, C.-K., Neylon, C., Hosking, R., Montgomery, L., Wilson, K.,
Ozaygen, A., \& Brookes-Kenworthy, C. (2020). \emph{Evaluating
institutional open access performance: Sensitivity analysis}.
\url{https://doi.org/10.1101/2020.03.19.998542}

@@

\bibitem[\citeproctext]{ref-Jahn_2025}
Jahn, N. (2025). How open are hybrid journals included in transformative
agreements? \emph{Quantitative Science Studies}, 1--39.
\url{https://doi.org/10.1162/qss_a_00348}

@@

\bibitem[\citeproctext]{ref-de_Jonge_2025}
Jonge, H. de, Kramer, B., \& Sondervan, J. (2025). \emph{Tracking
transformative agreements through open metadata: Method and validation
using dutch research council NWO funded papers.}
\url{https://doi.org/10.31222/osf.io/tz6be_v1}

@@

@@

@@

@@

\bibitem[\citeproctext]{ref-Rabesandratana_2019}
Rabesandratana, T. (2019). Elsevier deal with france disappoints
open-access advocates. \emph{Science}.
\url{https://doi.org/10.1126/science.aba5656}

@@

\bibitem[\citeproctext]{ref-Simard_2024}
Simard, M.-A., Basson, I., Hare, M., Lariviere, V., \& Mongeon, P.
(2024). \emph{The open access coverage of OpenAlex, scopus and web of
science}. \url{https://doi.org/10.48550/ARXIV.2404.01985}

@@

\bibitem[\citeproctext]{ref-Stahlschmidt_2022}
Stahlschmidt, S., \& Stephen, D. (2022). From indexation policies
through citation networks to normalized citation impacts: Web of
science, scopus, and dimensions as varying resonance chambers.
\emph{Scientometrics}, \emph{127}(5), 2413--2431.
\url{https://doi.org/10.1007/s11192-022-04309-6}

>>>>

Transformative agreements are a much-discussed library licensing model
to transition subscription-based journal publishing to full open access.
While these agreements can vary substantially, they mainly target hybrid
journal portfolios, enabling researchers from participating institutions
to publish open access in them while also providing reading access to
the entire portfolio. Responding to experiences with the subscription
model, library consortia and funders have begun to coordinate the
standardisation of negotiation, administration, and evaluation practices
for these agreements. Simultaneously, efforts to improve transparency
through open data about contracts and payments are ongoing. However,
this information remains only partially available through bibliometric
databases, limiting large-scale studies about the impact of
transformative agreements on the transition of hybrid journals to full
open access. This study compares findings from hoaddata, an openly
available, continuously updated dataset on hybrid open access, built on
open data from Crossref, OpenAlex and the cOAlition S Journal Checker
Tool, with the established bibliometric databases Web of Science and
Scopus. The purpose of this comparison is not only to evaluate the
suitability of open metadata for measuring hybrid open access but also
to inform bibliometric research and practice by highlighting the
strengths and weaknesses of each data source.

Since the initial proposal to further develop hybrid open access as a
means to cost-effectively repurpose subscription spending toward open
access (Schimmer et al., 2015), transformative agreements have led to
considerable academic debate. While standardisation activities,
particularly through the OA2020 and ESAC (Efficiency and Standards for
Article Charges) initiatives, have led to improved workflows between
library consortia and publishers, as well as increased open access
publishing (Campbell et al., 2022; Geschuhn \& Stone, 2017), the initial
focus on a few large publishers by national consortia from high-income
countries has drawn criticism (Borrego et al., 2021). Critics have in
particular raised concerns about market concentration (Butler et al.,
2023; Shu \& Larivière, 2023), the failure of most hybrid journals to
convert to full OA (Kiley, 2024; Momeni et al., 2021), reduced
competition and incentives (McCabe \& Mueller-Langer, 2024; Schmal,
2024), and the continuation of publication fees (APCs) that widen gaps
between well-resourced and under-resourced institutions and researchers,
raising equity concerns (Babini et al., 2022; Ross-Hellauer et al.,
2022).

In terms of bibliometric evidence, studying over 700 agreements, Jahn
(2025) observed a strong growth in open access publishing to 15\% in
2022 due to the varying implementation of transformative agreements
across countries, predominantly driven by European countries with
national consortia that successfully negotiated these agreements and the
three commercial publishers Elsevier, Springer Nature, and Wiley. This
aligns with the rapid proliferation of transformative agreements in
recent years (McCabe \& Mueller-Langer, 2024). Before, only a small
proportion of articles were made open access in hybrid journals Piwowar
et al. (2018), with only few European countries with dedicated hybrid
open access funding policies and the high-energy physics SCOAP\^{}3
consortium standing out (Huang et al., 2020; Kohls \& Mele, 2018;
Pinfield, 2015; Robinson-Garcia et al., 2020).

While consortia evaluations confirm the growth in open access publishing
through transformative agreements, they reach mixed policy conclusions.
For instance, the British Joint Information Systems Committee's
evaluation (Brayman et al., 2024) noted transformative agreements'
significant impact on national open access growth but limited effect on
global open access transition, suggesting a need to reevaluate their
strategy. Similar policy reconsiderations emerged in Norway and Sweden
(Holden et al., 2023; Widding, 2024). Meanwhile, cOAlition S recommended
ending financial support for these agreements in 2024, but still
considers hybrid open access compliant with funder open access policies.
Conversely, Germany's DEAL consortium extended agreements with major
publishers through 2028, while Colombia established Latin America's
first transformative agreements (Muñoz-Vélez et al., 2024). Presenting
experiences at the Max Planck Digital Library, (Dér, 2025) outlines
strategies to achieve full open access while controlling costs.

Open research information can benefit from transformative agreements as
library consortia also negotiate for more comprehensive open metadata to
avoid commodification of data analytics capabilities needed to negotiate
and evaluate agreements (Aspesi \& Brand, 2020; McCabe \&
Mueller-Langer, 2024). For instance, ESAC recommends using Crossref
(Hendricks et al., 2020) to share metadata supporting workflows
including open access licenses and funding information (Geschuhn \&
Stone, 2017). Simultaneously, the bibliometric community pushes for more
open metadata through this publisher-driven DOI registration platform
and continuously monitors progress (N. J. van Eck \& Waltman, 2022).

Numerous open scholarly data services building upon Crossref benefit
from these developments. A prominent example is OpenAlex, which has
gained considerable attention by promising inclusivity, openness, and
comprehensive coverage of publication metadata (Priem et al., 2022). In
a recent evaluation of OpenAlex, (Céspedes et al., 2025) reviewed the
current evidence base about its strengths and shortcomings. They
highlight both its increasing use in bibliometric research and adoption
across research organisations. Comparative studies generally perceive
OpenAlex as a reliable alternative to commercial databases (Alperin et
al., 2024; Culbert et al., 2024), although data quality issues were
observed regarding important metadata fields needed to analyse open
access, including evidence about open access status itself (Jahn et al.,
2023; Simard et al., 2024), as well as authors (Culbert et al., 2024),
affiliations (Zhang et al., 2024), author roles (Jahn, 2025), and
document types used to distinguish between different article types
(Haupka et al., 2024). However, these studies emphasise the new
possibilities of combining open metadata and OpenAlex's curation efforts
in conjunction with growing awareness about the importance of open
metadata as demonstrated by the Barcelona Declaration on Open Research
Information.

Against this background, this study evaluates the suitability of open
metadata to measure the impact of transformative agreements on hybrid
open access. Few studies have attempted this analysis due to data
limitations and the need to combine multiple sources. Jonge et al.
(2025) provides a notable exception by comparing transformative
agreement data from the cOAlition S Journal Checker tool and OpenAlex
with institutional invoicing records for publications funded by Dutch
research funders through the Dutch library consortium. Their analysis
found 89\% accuracy in identifying open access articles enabled by
transformative agreements, with discrepancies resulting from both open
data limitations and the inherent complexity of agreement structures
that require data from both library consortia and
publishers---information that is rarely made publicly available. Recent
large-scale studies (Bakker et al., 2024; Haustein et al., 2024; Jahn,
2025; Kramer, 2024) confirm that these challenges extend beyond the
Dutch context, highlighting monitoring difficulties related to varying
journal portfolios, challenges in using author roles as a proxy for
invoicing data, and article caps within agreements.

The present study addresses these gaps by adapting the open approach
used in Jahn (2025), which combines cOAlition S Journal Checker tool
data, Crossref and OpenAlex to measure the impact of transformative
agreements on the open access uptake of hybrid journals, and applying
this same methodology to Web of Science and Scopus data. This
comparative approach enables assessment of different bibliometric data
sources for monitoring transformative agreements while evaluating their
respective strengths and limitations. Specifically, this study compares
the indexing coverage of more than 13,000 hybrid journals across these
sources and examines whether consistent results can be obtained
regarding open access uptake between 2019 and 2023, analysed by
publisher and country. By comparing findings across these platforms,
this large-scale study aims to provide a more comprehensive
understanding of how transformative agreements impact the ongoing
transition from subscription-based journal publishing to full open
access.

@@

Throughout this largely automated data collection and analysis process,
Tidyverse tools (Wickham et al., 2019) for the R programming language (R
Core Team, 2024) were used. Rank correlations were calculated using the
Hmisc package (Harrell Jr, 2003). The source code analysis, including
all queries used to obtain the data, is available on GitHub (add link).
An interactive supplement exploring correlations between the data
sources examined by country and publisher is available on HuggingFace
Spaces: https://huggingface.co/spaces/najkoja/hoa\_replication.

@@

This study of over 13,000 hybrid journals shows a rise in open access
due to transformative agreements between 2019 and 2023, although most
articles remained paywalled. While transformative agreements accounted
for the majority of open access, many articles continue to become open
through the payment of individual publication fees. Hybrid open access
and transformative agreements remain concentrated among a small group of
large commercial publishers, with high-income European
countries---alongside South Africa---showing high adoption rates. In
contrast, the three most productive countries, China, the United States,
and India, were substantially ranked lower in transformative agreement
adoption. Open questions remain whether this uneven distribution
reflects temporary implementation gaps, inherent inequities in the
transformative agreement model, or a deliberate avoidance of such
agreements.

The findings were consistent across the investigated open data source
hoaddata, derived from Crossref and OpenAlex, and the established
proprietary bibliometric databases Scopus and Web of Science. Notably,
and aligning with previous studies (Akbaritabar et al., 2024; Alperin et
al., 2024) and rankings (N. J. van Eck et al., 2024), the results show
strong correlations by country affiliation, which supports using open
metadata for large-scale analyses of hybrid open access. However, the
observed differences in journal coverage and metadata availability
warrant discussion, affecting not only open data sources but also
proprietary databases when used in isolation.

The coverage analysis reveals that hybrid journals are well indexed in
all three data sources, particularly in terms of article coverage,
reflecting the dominance of major publishers whose established journal
portfolios are comprehensively indexed in proprietary databases (Bellen
et al., 2024). Differences emerge for journals targeting practitioners
or local non-English language communities, with many such titles indexed
exclusively in Crossref and Scopus. Using Crossref as a bibliometric
database in hoaddata demonstrated particular strength in identifying
newly established hybrid journals, a notable finding given that
transformative agreements primarily target existing subscription-based
journals. The landscape of hybrid journal publishing thus differs
markedly from that of fully open access journals. Comparing the coverage
of OpenAlex, Scopus and Web of Science, Simard et al. (2024) indicate
that only half of the fully open access journals listed in the DOAJ are
also indexed in Scopus and Web of Science. Notably, journals that charge
no publication fees (``diamond journals'') are absent from the selective
Web of Science, which reinforces existing disparities in the indexing of
under-represented research communities and regions in selective
bibliometric databases (Simard et al., 2024)

A frequently reported limitation of studying open access with less
selective databases is the lack of corresponding authorship information
(Fraser et al., 2023; Haucap et al., 2021; Shu \& Larivière, 2023).
However, this analysis demonstrates that indicators based on first
authors, which have often been used as a proxy for determining open
access funding, and corresponding authors show a high level of
correlation, reflecting disciplinary norms in scholarly publishing with
regard to contributorship and author roles and positions (Larivière et
al., 2016). First authors typically conduct the main research underlying
a paper, while the corresponding author often supervises the research
(Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly, measures
based on first or corresponding authorships are strongly correlated,
suggesting that these authors share the same country affiliation.
Moreover, in most cases, the first author is identical to the
corresponding author (Chinchilla-Rodríguez et al., 2024). Despite this
correlation, the study observed a recent slight increase in open access
articles by corresponding authors over first authors across Springer
Nature hybrid journals prompts questions about how institutional open
access sponsorship practices influence author roles and assignments
within co-author teams, especially as funding opportunities vary
(Gumpenberger et al., 2018). Previous research has highlighted how
institutionalized bibliometric practices can affect the valuation of
authorship positions (Helgesson, 2020), suggesting the need to monitor
potential influences of the availability of open access funding on
authorships (Maddi \& Silva, 2024).

@@

The database comparison revealed important discrepancies in open access
evidence. Crucially, not all publishers deposit CC licences with
Crossref, a limitation that becomes apparent when contrasting Crossref
with Unpaywall's data in Scopus and Web of Science. While Unpaywall can
detect such gaps by parsing publishers' websites for open access
licences, it nevertheless missed a substantial number of CC-licensed
open access articles from Wiley journals indexed in Crossref, likely due
to parsing errors on the publisher's website. This resulted in fewer
open access articles being recorded in Scopus and Web of Science.
Further inconsistencies emerged between Scopus and Web of Science,
despite both relying on Unpaywall: Web of Science erroneously labels
Elsevier's delayed open access as hybrid, whereas Scopus correctly
categorises it. Notably, Scopus and Crossref showed greater alignment
than observed in a related comparison between Scopus and OpenAlex
(Alperin et al., 2024). This likely reflects the use of a curated list
to identify hybrid open access journals, rather than relying solely on
article-level tags. Given these variations, which may compromise
comparability, research and monitoring exercises should avoid reliance
on a single source. Instead, the selection of open access evidence
should be cross-verified using multiple sources and snapshots that can
be used to track changes in the data over time (Huang et al., 2020).
Incorporating expert-curated journal lists can also help prevent
misclassification based on business models (Visser et al., 2021).

It is important to note that the study's estimates of articles from
institutions involved in transformative agreements are approximations
due to a lack of access to invoice data, which is not usually shared by
library consortia and publishers. Furthermore, the data sources about
transformative agreements including the cOAlition S Journal Checker Tool
and the ESAC registry are a voluntary effort, crowd-sourced from various
consortia. However, recent validation with Dutch research information
demonstrates the reliability of such an open approach for the assessment
of articles under transformative agreements (Jonge et al., 2025). With
enhanced open metadata compliance, attributable to evolving standards
and initiatives to support negotiations with publishers, particularly
through the communities of the ESAC initiative and Barcelona Declaration
supporters, the situation is likely to improve. In conclusion, this
study has shown that open metadata are suitable for analysing the
transition from subscription-based journal publishing to full open
access.

@@

\bibitem[\citeproctext]{ref-Aspesi_2020}
Aspesi, C., \& Brand, A. (2020). In pursuit of open science, open access
is not enough. \emph{Science}, \emph{368}(6491), 574--577.
\url{https://doi.org/10.1126/science.aba3763}

@@

\bibitem[\citeproctext]{ref-budapest}
Babini, D., Chan, L., Hagemann, M., Joseph, H., Kuchma, I., \& Suber, P.
(2022). \emph{{The Budapest Open Access Initiative-20th. Anniversary
recommendations (BOAI20)}}.
\url{https://www.budapestopenaccessinitiative.org/boai20/}

\bibitem[\citeproctext]{ref-Bakker_2024}
Bakker, C., Langham-Putrow, A., \& Riegelman, A. (2024). Impact of
transformative agreements on publication patterns: An analysis based on
agreements from the ESAC registry. \emph{International Journal of
Librarianship}, \emph{8}(4), 67--96.
\url{https://doi.org/10.23974/ijol.2024.vol8.4.341}

\bibitem[\citeproctext]{ref-vanbellen2024oligopolyacademicpublisherspersists}
Bellen, S. van, Alperin, J. P., \& Larivière, V. (2024). \emph{The
oligopoly of academic publishers persists in exclusive database}.
\url{https://arxiv.org/abs/2406.17893}

\bibitem[\citeproctext]{ref-Birkle_2020}
Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). {Web of
Science} as a data source for research on scientific and scholarly
activity. \emph{Quantitative Science Studies}, \emph{1}(1), 363--376.
\url{https://doi.org/10.1162/qss_a_00018}

@@

\bibitem[\citeproctext]{ref-Jisc_2024}
Brayman, K., Devenney, A., Dobson, H., Marques, M., \& Vernon, A.
(2024). \emph{A review of transitional agreements in the {UK}}. Zenodo.
\url{https://doi.org/10.5281/zenodo.10787392}

\bibitem[\citeproctext]{ref-Butler_2023}
Butler, L.-A., Matthias, L., Simard, M.-A., Mongeon, P., \& Haustein, S.
(2023). The oligopoly's shift to open access: How the big five academic
publishers profit from article processing charges. \emph{Quantitative
Science Studies}, \emph{4}(4), 778--799.
\url{https://doi.org/10.1162/qss_a_00272}

\bibitem[\citeproctext]{ref-ifla}
Campbell, C., Dér, Á., Geschuhn, K., \& Valente, A. (2022). How are
transformative agreements transforming libraries? \emph{87th IFLA World
Library and Information Congress (WLIC) / 2022 in Dublin, Ireland}.
\url{https://repository.ifla.org/handle/123456789/1973}

\bibitem[\citeproctext]{ref-C_spedes_2025}
Céspedes, L., Kozlowski, D., Pradier, C., Sainte‐Marie, M. H., Shokida,
N. S., Benz, P., Poitras, C., Ninkov, A. B., Ebrahimy, S., Ayeni, P.,
Filali, S., Li, B., \& Larivière, V. (2025). Evaluating the linguistic
coverage of {OpenAlex}: An assessment of metadata accuracy and
completeness. \emph{Journal of the Association for Information Science
and Technology}. \url{https://doi.org/10.1002/asi.24979}

@@

\bibitem[\citeproctext]{ref-culbert2024referencecoverageanalysisopenalex}
Culbert, J., Hobert, A., Jahn, N., Haupka, N., Schmidt, M., Donner, P.,
\& Mayr, P. (2024). \emph{Reference coverage analysis of OpenAlex
compared to web of science and scopus}.
\url{https://arxiv.org/abs/2401.16359}

\bibitem[\citeproctext]{ref-D_r_2025}
Dér, Á. (2025). What gets missed in the discourse on transformative
agreements. \emph{Katina Magazine}.
\url{https://doi.org/10.1146/katina-20250212-1}

@@

\bibitem[\citeproctext]{ref-eck_crossref_2022}
Eck, N. J. van, \& Waltman, L. (2022). \emph{Crossref as a source of
open bibliographic metadata}. MetaArXiv.
\url{https://doi.org/10.31222/osf.io/smxe5}

@@

\bibitem[\citeproctext]{ref-Harrell_Jr_2003}
Harrell Jr, F. E. (2003). Hmisc: Harrell miscellaneous. In \emph{CRAN:
Contributed Packages}. The R Foundation.
\url{https://doi.org/10.32614/cran.package.hmisc}

@@

\bibitem[\citeproctext]{ref-norway}
Holden, L., Skoie, M., Røeggen, V., Bjerde, K. W., Wenaas, L., Bakke,
P., Løvhaug, J. W., Karlsen, E. S., \& Qvenild, M. (2023).
\emph{Strategi for vitenskapelig publisering etter 2024}. Sikt.
\url{https://doi.org/10.18711/2KZ1-BA97}

@@

\bibitem[\citeproctext]{ref-Huang_2020}
Huang, C.-K., Neylon, C., Hosking, R., Montgomery, L., Wilson, K.,
Ozaygen, A., \& Brookes-Kenworthy, C. (2020). \emph{Evaluating
institutional open access performance: Sensitivity analysis}. bioRxiv.
\url{https://doi.org/10.1101/2020.03.19.998542}

@@

\bibitem[\citeproctext]{ref-Jahn_2025}
Jahn, N. (2025). How open are hybrid journals included in transformative
agreements? \emph{Quantitative Science Studies}, \emph{6}, 242--262.
\url{https://doi.org/10.1162/qss_a_00348}

\bibitem[\citeproctext]{ref-jahn2023analysing}
Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
reclassifying open access information in OpenAlex}. {Scholarly
Communication Analytics Blog}.
\url{https://subugoe.github.io/scholcomm_analytics/posts/oalex_oa_status/}

@@

\bibitem[\citeproctext]{ref-de_Jonge_2025}
Jonge, H. de, Kramer, B., \& Sondervan, J. (2025). \emph{Tracking
transformative agreements through open metadata: Method and validation
using {Dutch Research Council NWO} funded papers}. MetaArXiv.
\url{https://doi.org/10.31222/osf.io/tz6be_v1}

\bibitem[\citeproctext]{ref-Kiley_2024}
Kiley, R. (2024). \emph{{Transformative Journals}: Analysis from the
2023 reports}. {cOAlition S}.
\url{https://www.coalition-s.org/blog/transformative-journals-analysis-from-the-2023-reports/}

\bibitem[\citeproctext]{ref-Kohls_2018}
Kohls, A., \& Mele, S. (2018). Converting the literature of a scientific
field to open access through global collaboration: The experience of
SCOAP3 in particle physics. \emph{Publications}, \emph{6}(2), 15.
\url{https://doi.org/10.3390/publications6020015}

\bibitem[\citeproctext]{ref-Kramer_2024}
Kramer, B. (2024). \emph{Study on scientific publishing in {Europe} --
{Development}, diversity, and transparency of costs}. Publications
Office of the European Union. \url{https://doi.org/doi/10.2777/89349}

@@

\bibitem[\citeproctext]{ref-Laakso_2016}
Laakso, M., \& Björk, B.-C. (2016). Hybrid open access--a longitudinal
study. \emph{Journal of Informetrics}, \emph{10}(4), 919--932.
\url{https://doi.org/10.1016/j.joi.2016.08.002}

@@

\bibitem[\citeproctext]{ref-McCabe_2024}
McCabe, M. J., \& Mueller-Langer, F. (2024). Open access is shaping
scientific communication. \emph{Science}, \emph{385}(6714), 1170--1172.
\url{https://doi.org/10.1126/science.adp8882}

@@

\bibitem[\citeproctext]{ref-Momeni_2021}
Momeni, F., Mayr, P., Fraser, N., \& Peters, I. (2021). What happens
when a journal converts to open access? A bibliometric analysis.
\emph{Scientometrics}, \emph{126}(12), 9811--9827.
\url{https://doi.org/10.1007/s11192-021-03972-5}

\bibitem[\citeproctext]{ref-Mu_oz_V_lez_2024}
Muñoz-Vélez, H., Pallares, C., Echavarría, A. F., Contreras, J., Pavas,
A., Bello, D., Rendón, C., Calderón-Rojas, J., \& Garzón, F. (2024).
Strategies for negotiating and signing transformative agreements in the
{Global South}: The {Colombia Consortium} experience. \emph{Journal of
Library Administration}, \emph{64}(1), 80--98.
\url{https://doi.org/10.1080/01930826.2023.2287945}

\bibitem[\citeproctext]{ref-Pinfield_2015}
Pinfield, S. (2015). Making open access work: The {``state-of-the-art''}
in providing open access to scholarly literature. \emph{Online
Information Review}, \emph{39}(5), 604--636.
\url{https://doi.org/10.1108/oir-05-2015-0167}

@@

\bibitem[\citeproctext]{ref-r}
R Core Team. (2024). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-Rabesandratana_2019}
Rabesandratana, T. (2019). Elsevier deal with {France} disappoints
open-access advocates. \emph{Science}.
\url{https://doi.org/10.1126/science.aba5656}

\bibitem[\citeproctext]{ref-Robinson_Garcia_2020}
Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
access uptake by universities worldwide. \emph{{PeerJ}}, \emph{8},
e9410. \url{https://doi.org/10.7717/peerj.9410}

\bibitem[\citeproctext]{ref-Ross_Hellauer_2022}
Ross-Hellauer, T., Reichmann, S., Cole, N. L., Fessl, A., Klebel, T., \&
Pontika, N. (2022). Dynamics of cumulative advantage and threats to
equity in open science: A scoping review. \emph{Royal Society Open
Science}, \emph{9}(1). \url{https://doi.org/10.1098/rsos.211032}

\bibitem[\citeproctext]{ref-Schimmer_2015}
Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{{Disrupting the
subscription journals'business model for the necessary large-scale
transformation to open access}}. Max Planck Digital Library.
\url{https://doi.org/10.17617/1.3}

\bibitem[\citeproctext]{ref-Schmal_competition}
Schmal, W. B. (2024). \emph{La révolution dévore ses enfants: Pricing
implications of transformative agreements}. arXiv.
\url{https://doi.org/10.48550/ARXIV.2403.03597}

@@

\bibitem[\citeproctext]{ref-Simard_2024}
Simard, M.-A., Basson, I., Hare, M., Lariviere, V., \& Mongeon, P.
(2024). \emph{The open access coverage of {OpenAlex}, {Scopus} and {Web
of Science}}. \url{https://doi.org/10.48550/ARXIV.2404.01985}

@@

\bibitem[\citeproctext]{ref-Stahlschmidt_2022}
Stahlschmidt, S., \& Stephen, D. (2022). From indexation policies
through citation networks to normalized citation impacts: {Web of
Science}, {Scopus}, and {Dimensions} as varying resonance chambers.
\emph{Scientometrics}, \emph{127}(5), 2413--2431.
\url{https://doi.org/10.1007/s11192-022-04309-6}
